Inferring Continuous and Discrete Population Genetic Structure Across Space

### 2017-08-17 writeup/conStruct.tex
moved allelic cov to Appendix

%%%%%%%%%%%%%%%%%
\section*{Methods}

@@

In addition to these differences on the diagonal,
this differs from the usual ``genetic covariance'' \citep{mcvean_genealogical_2009}
in that (a) we do not subtract locus means,
and so to make the statistic invariant under choices of reference allele, 
(b) we randomly choose alleles to count.
To see how to arrive at equation \eqref{allelic_covariance},
pick a random locus and
let $A$ and $B$ be randomly drawn alleles at that locus from populations $i$ and $j$ respectively.
If $i=j$ then these are chosen \emph{without} replacement.
Suppose these are each coded as `0' or `1` (where `0' denotes a reference allele),
but we randomly ``flip'' this coding, so that we let $X=A$ and $Y=B$ with probability 1/2,
but otherwise we let $X=1-A$ and $Y=1-B$.
We then let $\widehat{\Omega}_{i,j} = \cov[X,Y]$ --
sampling without replacement makes the statistic insensitive to sample size,
and random allele flipping makes it independent of the choice of reference allele.
By conditioning on the flip,
and using the fact that $\E[X] = \E[Y] = 1/2$,
equation \eqref{allelic_covariance} comes from the calculation that
$\cov[X,Y] = \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4 = \E[(A-1/2)(B-1/2)]$.
The factor of $\delta_{i,j}/(N_i-1)$ comes from the fact that if $i=j$,
then $\E[AB]$ is the probability that both alleles drawn without replacement are reference,
which is $f_i^2 n/(n-1)$.
As noted in \cite{EEMS},
this covariance has a close relationship to the pairwise genetic distance:
if $D_{i,j}$ is the mean density of sites at which random samples from $i$ and $j$ differ at a randomly chosen locus,
then $\Sigma_{i,j} = (1 - 2 D_{i,j})/4$.
Therefore, unlike genetic covariance,
this allelic covariance is affected in shape by singleton sites,
so it may be advisable to filter these prior to analysis
if these are likely to contain a large percentage of errors.

@@

\begin{figure}
	\centering
		{\includegraphics[width=4in,height=2.5in]{figs/schematic/method_schematic.png}}
		\caption{Schematic of our method, using $K=3$ as an example.
			     The pattern of isolation by distance within each layer is denoted by the color gradient in each of the layer ellipses, 
			     and $\phi_{(k)}$ denotes the covariance shared by samples with ancestry entirely in the $k$th cluster.
			     Sampled populations on the landscape are inferred to be admixed between these layers; 
			     the $i$th sample draws proportion $w^{(k)}_i$ of its ancestry from cluster $k$.
			    }\label{sim_setup}
\end{figure}

@@

A nice feature of our approach is that the model described in Eqn \refeq{cross_cluster_covariance} 
contains a nonspatial assignment model as a special case 
(see \ref{model_app} for a more in-depth discussion). 
By setting $\alpha_0^{(k)}$ to zero for all $k$, 
we obtain a nonspatial model in which each cluster has its own allele frequency at each SNP, 
and individuals draw a proportion of their ancestry from each cluster. 
This model is strongly conceptually similar to that of STRUCTURE \citep{STRUCTURE} 
and related models \citep[e.g.][]{ADMIXTURE}; 
the only difference is that our likelihood assumes that allele frequencies are normally distributed 
around their expectations, 
while the standard assignment methods assume that the error is binomial distributed (see XXXX).  
Therefore, we can compare the fit of the different models 
--- spatial vs. nonspatial, across different values of $K$ --- 
by comparing their performance in a common framework.

@@

\gc{IS PREDICTION accuracy the right phrase?}
We choose to take a cross-validation approach \citep[similar in spirit to][]{ADMIXTURE_xval}.
We estimate the posterior on the parameters of the model using a ``training" partition of the data 
(in practice the training set is a random subset of $90\%$ of the loci). 
We then calculate the likelihood of the ``testing" subset (the remaining $10\%$ of the loci), 
averaged over the posterior of the parameters. 
We repeat this procedure a number of times, 
take an average of the relative log-likelihoods over the runs at each $K$, 
and select the model with the significantly best predictive accuracy
(see Section \ref{Xvalidation} for more on our cross-validation procedure).
This approach rewards models with larger values of $K$, 
which increase the likelihood of the training partition, 
as they fit the data better,
but the improvement in the likelihood will plateau (or peak), 
as above a certain $K$ the model is fitting noise specific to the training data 
and therefore not fitting the testing set any better.
\plr{I wonder if we can just cite \cite{verity_nichols2016} rather than this laundry list here?}

We use cross-validation to attack this problem.
To do this,
we use a ``training" partition of the data (in practice, a random 90\% subset of the loci)
to estimate the posterior distribution of the parameters,
and then calculate the log-likelihood of the remaining ``testing'' loci,
averaged over the posterior.
Prediction accuracy of a particular model (e.g., value of $K$)
is then measured using this log-likelihood,
averaged over a number of independent data partitions.
The best model is judged to be the simplest one with significantly better predictive accuracy
than others (see Section \ref{Xvalidation} for more on our cross-validation procedure).
In general, larger values of $K$ allows the model more flexibility,
and thus increases the likelihood of the training partition, 
but this improvement in the likelihood will plateau (or even peak), 
since above a certain $K$ the model only fits noise specific to the training data 
rather than generalizable patterns.
At any value of $K$, support for the spatial model over the nonspatial model 
means that isolation by distance is likely a feature of the data.

@@

\subsection*{Simulations}
We first test the method with scenarios generated using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
To generate discrete population structure, we simulate $K$ distinct populations,
each of which split from a common ancestor $\tau_{\text{s}}$ units of coalescent time in the past,
without subsequent migration between them.
To generate continuous differentiation within each cluster,
at time $\tau_{\text{e}}$ in the past,
each of these discrete clusters instantaneously colonizes a lattice of demes,
for which we use a stepping stone model with symmetric migration 
to nearest neighbors (eight neighbors, including diagonals).
Finally, at time $\tau_{\text{a}}$ in the past 
we generate a single dataset % that captures both continuous and discrete patterns of population structure 
by collapsing those $K$ discrete lattices into a single grid of demes that are admixed to various degrees from these different layers.
%(see Fig \ref{sim_setup}).
We simulated datasets using $K=1$, 2, and 3 layers,
in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.

%\begin{figure}
%	\centering
%		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
%		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
%			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
%			    then each expand to colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
%			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
%			    with ancestry entirely in one or the other of the clusters,
%			    or admixed between them.
%			    }\label{sim_setup}
%\end{figure}

We ran \texttt{conStruct} analyses on each of these simulated datasets, 
specifying both a spatial and nonspatial model for $K = $ 1 through 7, 
then used cross-validation to compare the predictive performance of the models.
In Fig \ref{simK1_pies}, we show the results of the nonspatial model 
run with between 1 and 7 clusters, applied to the dataset simulated under $K=1$,
and in Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
The admixture pie plots for the spatial and nonspatial models run on the datasets simulated under $K=2$ and $K=3$
are shown in Figs \ref{simK1_nsp_pies}-\ref{simK3_sp_pies}.
To confirm that results from our nonspatial model were qualitatively similar to those of 
other nonspatial model-based clustering algorithms, 
we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
inferring admixture under models with $K=2$ through 4 (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).

A number of results are immediately clear from these tests with simulated data.
First, \texttt{conStruct} performs well on all simulated scenarios.
Across all values of K and for all simulations, 
the spatial model is correctly preferred over the nonspatial model, 
and it infers the true admixture proportions with high accuracy, tight precision, and good coverage 
(Fig \ref{sim_xvals}).
Second, the predictive accuracy of the nonspatial model increases 
with the number of clusters in the model, 
as it describes continuous, spatial patterns of relatedness between demes within the same cluster 
using gradients of admixture across more and more discrete clusters.
This behavior of the nonspatial model is highlighted in 
the analyses of data simulated under $K=1$ (Fig \ref{simK1_pies} and \ref{simK1_nsp_pies}):
as $K$ in the model increases, 
the sampled area is increasingly divvied up between clusters 
with ``centers of mass" in the $K$ geographical extremes of the simulated space.
At $K=2$, the clusters are centered in opposite corners on the map; 
at $K=3$, in the lower two corners and the top half; 
at $K=4$, all four corners, and so on.
While it is true that each such cluster is more similar within itself
than it is to other clusters,
we know that these boundaries are arbitray,
since the data were simulated with no such boundaries.

@@

Trees in the genus \textit{Populus} (poplars, aspens, and cottonwoods) 
are distributed throughout the Northern Hemisphere, 
and include species of ecological and economic importance.
Species in the genus regularly co-occur and, 
where they do, they frequently hybridize \citep{eckenwalder1984, Cronk2005}.

Here, we focus on two \textit{Populus} species in the Pacific Northwest: 
\textit{Populus trichocarpa}, the black cottonwood,
and \textit{Populus balsamifera}, the balsam poplar.
These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
Both species are sampled over a large geographic region, 
and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
making the system well-suited to an analysis designed to describe
how genetic structure is partitioned both discretely across population (or species) clusters
and continuously with geographic distance.
We had the following questions for the system: 
\begin{enumerate}
\item Given the known history of hybridization between this pair of poplar species, 
is there support for two discrete clusters, 
rather than a single cline of ancestry across their overlap zone?
\item If there is evidence for discrete population structuring, 
does it fall solely along species boundaries, 
or is there sub-structuring between different populations of the same species?
\item Is there support for a pattern of isolation by distance within
the inferred clusters, and, if so, 
does it differ between different layers?
\end{enumerate}

@@

All models with $K>1$ assigned the majority of each species
to distinct layers, with some populations drawing ancestry from both layers.
This provides strong support for discrete population structure between the two species,
with some admixture,
% Although admixture was inferred between \textit{P. balsamifera} and \tri{}
% across all values of $K>1$, and in both the spatial and nonspatial models,
% there is nonetheless strong evidence for discrete population structuring, 
rather than a single, continuous cline of ancestry.
At all values of $K>1$, discrete population structure was mostly partitioned along species lines; 
at values of $K$ above 2, further discrete substructure was inferred within the \textit{P. trichocarpa} samples, 
while \bals{} samples displayed no substructure up to $K=7$.
This lack of substructure within the sampled \bals{} populations is in agreement with results from 
\citet{keller_etal_2010}, who found that \bals{} in the region is characterized by relatively low diversity, 
and probably represents a recent, postglacial expansion.

@@

\paragraph{Likelihood}
If our allele frequency deviations are well approximated by a Gaussian, 
their sample allelic covariance is a sufficient statistic,
so that calculating the likelihood of their sample allelic covariance is the same as 
calculating the probability of the frequency data up to a constant. 
We can therefore model the covariance of the sample allele frequencies, $\widehat{\Omega}$, 
as a draw from a Wishart distribution with degrees of freedom equal to 
the number of loci $L$ across which the sample allelic covariance is calculated:

\begin{align}
&\widehat{\Omega}_{i,j} = 
	\begin{cases}
		\frac{1}{L-1}
		\sum\limits_{\ell=1}^{L}{\left[ 	\left(f_{i,\ell} - \bar{f}_{i} \right)
		 	\left(f_{j,\ell} - \bar{f}_{j} \right)\right] \notag
			} & \text{for } i \neq j \\
		\bar{f}_i (1-\bar{f}_i) & \text{for } i = j \\
	\end{cases}\\
&\widehat{\Omega} \sim \mathcal{W}\left( L\Omega, L	\right)
\label{wishart}
\end{align}

@@

\section{Models, parameters, and priors}\label{model_app}
\paragraph{Spatial vs. nonspatial}
In this paper, we discuss two types of models, spatial and nonspatial, 
each of which can be implemented with different numbers of clusters.
The spatial model is parameterized as in Eqn \ref{admixed_spatial_cov},
and the nonspatial model is described in Eqn \ref{admixed_discrete_covariance}.
The nonspatial model is a special case of the spatial model, 
in which the value of all $\alpha$ parameters is set to 0.
The nonspatial model therefore has $3K$ fewer parameters than the spatial model,
as there are three $\alpha$ parameters that describe the continuous differentiation effect of distance per cluster,
and they are fixed in the nonspatial model.

@@

\section{Our Model vs. STRUCTURE}
We should probably demonstrate that the nonspatial version of \texttt{conStruct} is basically the same as STRUCTURE.
Maybe check out Engelhardt \& Stephens?

@@

>>>>

\section*{Methods}

@@

In addition to these differences on the diagonal,
this differs from the usual ``genetic covariance'' \citep{mcvean_genealogical_2009}
in that (a) we do not subtract locus means,
and so to make the statistic invariant under choices of reference allele, 
(b) we randomly choose an allele to count at each locus.
We discuss the derivation of Eqn. \eqref{allelic_covariance} further in the Appendix (\secref{allelic_cov}).
As noted in \cite{EEMS},
this covariance has a close relationship to the pairwise genetic distance:
if $D_{i,j}$ is the mean density of sites at which random samples from $i$ and $j$ differ at a randomly chosen locus,
then $\Sigma_{i,j} = (1 - 2 D_{i,j})/4$.
Therefore, unlike genetic covariance,
this allelic covariance is affected in shape by singleton sites,
so it may be advisable to filter these prior to analysis
if these are likely to contain a large percentage of errors.

@@

\begin{figure}
	\centering
		{\includegraphics[width=4in,height=2.5in]{figs/schematic/method_schematic.png}}
		\caption{Schematic of our method, using $K=3$ as an example.
			     The pattern of isolation by distance within each layer is denoted by the color gradient in each of the layer ellipses, 
			     and $\phi_{(k)}$ denotes the covariance shared by samples with ancestry entirely in the $k$th cluster.
			     Sampled populations on the landscape are inferred to be admixed between these layers; 
			     the $i$th sample draws proportion $w^{(k)}_i$ of its ancestry from cluster $k$.
			    }\label{schematic}
\end{figure}

@@

A nice feature of our approach is that the model described in Eqn \eqref{cross_cluster_covariance} 
contains a nonspatial assignment model as a special case 
(see \secref{model_app} for a more in-depth discussion). 
By setting $\alpha_0^{(k)}$ to zero for all $k$, 
we obtain a nonspatial model in which each cluster has its own allele frequency at each SNP, 
and individuals draw a proportion of their ancestry from each cluster. 
This model is strongly conceptually similar to that of STRUCTURE \citep{STRUCTURE} 
and related models \citep[e.g.][]{ADMIXTURE}; 
the only difference is that our likelihood assumes that allele frequencies are normally distributed 
around their expectations, 
while the standard assignment methods assume that the error is binomial distributed (see XXXX).  
Therefore, we can compare the fit of the different models 
--- spatial vs. nonspatial, across different values of $K$ --- 
by comparing their performance in a common framework.

@@

We choose to take a cross-validation approach \citep[similar in spirit to][]{ADMIXTURE_xval}.
We estimate the posterior on the parameters of the model using a ``training" partition of the data 
(in practice the training set is a random subset of $90\%$ of the loci). 
We then calculate the likelihood of the ``testing" subset (the remaining $10\%$ of the loci), 
averaged over the posterior of the parameters. 
We repeat this procedure a number of times, 
take an average of the relative log-likelihoods over the runs at each $K$, 
and select the model with the significantly best predictive accuracy
(see \secref{Xvalidation} for more on our cross-validation procedure).
This approach rewards models with larger values of $K$, 
which increase the likelihood of the training partition, 
as they fit the data better,
but the improvement in the likelihood will plateau (or peak), 
as above a certain $K$ the model is fitting noise specific to the training data 
and therefore not fitting the testing set any better.
\plr{I wonder if we can just cite \cite{verity_nichols2016} rather than this laundry list here?}

We use cross-validation to attack this problem.
To do this,
we use a ``training" partition of the data (in practice, a random 90\% subset of the loci)
to estimate the posterior distribution of the parameters,
and then calculate the log-likelihood of the remaining ``testing'' loci,
averaged over the posterior.
Prediction accuracy of a particular model (e.g., value of $K$)
is then measured using this log-likelihood,
averaged over a number of independent data partitions.
The best model is judged to be the simplest one with significantly better predictive accuracy
than others (see \secref{Xvalidation} for more on our cross-validation procedure).
In general, larger values of $K$ allows the model more flexibility,
and thus increases the likelihood of the training partition, 
but this improvement in the likelihood will plateau (or even peak), 
since above a certain $K$ the model only fits noise specific to the training data 
rather than generalizable patterns.
At any value of $K$, support for the spatial model over the nonspatial model 
means that isolation by distance is likely a feature of the data.

@@

\subsection*{Simulations}
We first test the method with scenarios generated using the coalescent simulator \texttt{ms} \citep{Hudson2002}.
To generate discrete population structure, we simulate $K$ distinct populations,
each of which split from a common ancestor $\tau_{\text{s}}$ units of coalescent time in the past,
without subsequent migration between them.
To generate continuous differentiation within each cluster,
at time $\tau_{\text{e}}$ in the past,
each of these discrete clusters instantaneously colonizes a lattice of demes,
for which we use a stepping stone model with symmetric migration 
to nearest neighbors (eight neighbors, including diagonals).
Finally, at time $\tau_{\text{a}}$ in the past 
we generate a single dataset 
by collapsing those $K$ discrete lattices into a single grid of demes that are admixed to various degrees from these different layers.
(see Fig \ref{sim_setup}).
We simulated datasets using $K=1$, 2, and 3 layers,
in each sampling 10,000 unlinked loci from each of 20 haploid individuals from every deme.

We ran \texttt{conStruct} analyses on each of these simulated datasets, 
specifying both a spatial and nonspatial model for $K = $ 1 through 7, 
then used cross-validation to compare the predictive performance of the models.
In Fig \ref{simK1_pies}, we show the results of the nonspatial model 
run with between 1 and 7 clusters, applied to the dataset simulated under $K=1$,
and in Fig \ref{sim_xvals} we show the results of the cross-validation analysis for each of the three simulated datasets.
The admixture pie plots for the spatial and nonspatial models run on the datasets simulated under $K=2$ and $K=3$
are shown in Figs \ref{simK1_nsp_pies}-\ref{simK3_sp_pies}.
To confirm that results from our nonspatial model were qualitatively similar to those of 
other nonspatial model-based clustering algorithms, 
we also analyzed each simulated dataset using fastSTRUCTURE \citep{fastStructure}, 
inferring admixture under models with $K=2$ through 4.

A number of results are immediately clear from these tests with simulated data.
First, \texttt{conStruct} performs well on all simulated scenarios.
Across all values of K and for all simulations, 
the spatial model is correctly preferred over the nonspatial model, 
and it infers the true admixture proportions with high accuracy, tight precision, and good coverage 
(Fig \ref{sim_xvals}).
Second, the predictive accuracy of the nonspatial model increases 
with the number of clusters in the model, 
as it describes continuous, spatial patterns of relatedness between demes within the same cluster 
using gradients of admixture across more and more discrete clusters.
This behavior of the nonspatial model is highlighted in 
the analyses of data simulated under $K=1$ (Fig \ref{simK1_pies} and \ref{simK1_nsp_pies}):
as $K$ in the model increases, 
the sampled area is increasingly divvied up between clusters 
with ``centers of mass" in the $K$ geographical extremes of the simulated space.
At $K=2$, the clusters are centered in opposite corners on the map; 
at $K=3$, in the lower two corners and the top half; 
at $K=4$, all four corners, and so on.
These results are qualitatively similar to those generated using fastSTRUCTURE \citep{fastStructure}, 
which also describes continuous differentiation using admixture between clusters
 (Fig \ref{fastStr_simK1}-\ref{fastStr_simK3}).
While it is true that each such cluster is more similar within itself
than it is to other clusters,
we know that these boundaries are arbitrary,
since the data were simulated with no such boundaries.

@@

Trees in the genus \textit{Populus} (poplars, aspens, and cottonwoods) 
are distributed throughout the Northern Hemisphere;
species in the genus regularly co-occur and, 
where they do, they frequently hybridize \citep{eckenwalder1984, Cronk2005}.

Here, we focus on two \textit{Populus} species in the Pacific Northwest: 
\textit{Populus trichocarpa}, the black cottonwood,
and \textit{Populus balsamifera}, the balsam poplar.
These species have a broad zone of overlap and are hypothesized to hybridize \citep{geraldes_etal_2014, suarezgonzalez_etal_2016}.
Both species are sampled over a large geographic region, 
and show spatial patterns of genetic and phenotypic variation \citep{slavov_etal_2012, mckown_etal_2013},
making the system well-suited to test an analysis designed to describe
how genetic structure is partitioned both discretely across population (or species) clusters
and continuously with geographic distance.
We organize the results of our analyses around the following questions: 
\begin{enumerate}
\item Given the known history of hybridization between this pair of poplar species, 
is there support for two discrete clusters, 
rather than a single cline of ancestry across their overlap zone?
\item If there is evidence for discrete population structuring, 
does it fall solely along species boundaries, 
or is there sub-structuring between different populations of the same species?
\item Is there support for a pattern of isolation by distance within
the inferred clusters, and, if so, 
does it differ between different layers?
\end{enumerate}

@@

All models with $K>1$ assigned the majority of each species
to distinct layers, with some populations drawing ancestry from both layers.
This provides strong support for discrete population structure between the two species,
with some admixture,
rather than a single, continuous cline of ancestry.
At all values of $K>1$, discrete population structure was mostly partitioned along species lines; 
at values of $K$ above 2, further discrete substructure was inferred within the \textit{P. trichocarpa} samples, 
while \bals{} samples displayed no substructure up to $K=7$.
This lack of substructure within the sampled \bals{} populations is in agreement with results from 
\citet{keller_etal_2010}, who found that \bals{} in the region is characterized by relatively low diversity, 
and probably represents a recent, postglacial expansion.

@@

\section{Allelic Covariance}\label{allelic_cov}
To see how to arrive at Eqn. \eqref{allelic_covariance},
pick a random locus and
let $A$ and $B$ be randomly drawn alleles at that locus from populations $i$ and $j$ respectively.
If $i=j$, then these are chosen \emph{without} replacement.
Suppose these are each coded as `0' or `1` (where `0' denotes a reference allele),
but we randomly ``flip'' this coding, so that we let $X=A$ and $Y=B$ with probability 1/2,
but otherwise we let $X=1-A$ and $Y=1-B$.
We then let $\widehat{\Omega}_{i,j} = \cov[X,Y]$ ; 
sampling without replacement makes the statistic insensitive to sample size,
and random allele flipping makes it independent of the choice of reference allele.
By conditioning on the flip,
and using the fact that $\E[X] = \E[Y] = 1/2$,
Eqn. \eqref{allelic_covariance} comes from the calculation that
\begin{align}
\cov[X,Y] &= \E[XY] - 1/4 = (1/2)\E[AB + (1-A)(1-B)] - 1/4\\\notag
&= \E[(A-1/2)(B-1/2)]
\end{align}
The factor of $\delta_{i,j}/(N_i-1)$ comes from the fact that if $i=j$,
then $\E[AB]$ is the probability that both alleles drawn without replacement are reference,
which is $f_i^2 n/(n-1)$.
\paragraph{Likelihood}
If our allele frequency deviations are well approximated by a Gaussian, 
their sample allelic covariance is a sufficient statistic,
so that calculating the likelihood of their sample allelic covariance is the same as 
calculating the probability of the frequency data up to a constant. 
We can therefore model the covariance of the sample allele frequencies, $\widehat{\Omega}$, 
as a draw from a Wishart distribution with degrees of freedom equal to 
the number of loci $L$ across which the sample allelic covariance is calculated:
\gb{should we take this equation out?}
\begin{align}
&\widehat{\Omega}_{i,j} = 
	\begin{cases}
		\frac{1}{L-1}
		\sum\limits_{\ell=1}^{L}{\left[ 	\left(f_{i,\ell} - \bar{f}_{i} \right)
		 	\left(f_{j,\ell} - \bar{f}_{j} \right)\right] \notag
			} & \text{for } i \neq j \\
		\bar{f}_i (1-\bar{f}_i) & \text{for } i = j \\
	\end{cases}\\
&\widehat{\Omega} \sim \mathcal{W}\left( L\Omega, L	\right)
\label{wishart}
\end{align}

@@

\section{Models, parameters, and priors}\label{model_app}
\paragraph{Spatial vs. nonspatial}
In this paper, we discuss two types of models, spatial and nonspatial, 
each of which can be implemented with different numbers of clusters.
The spatial model is parameterized as in Eqn \eqref{admixed_spatial_cov},
and the nonspatial model is described in Eqn \eqref{admixed_discrete_covariance}.
The nonspatial model is a special case of the spatial model, 
in which the value of all $\alpha$ parameters is set to 0.
The nonspatial model therefore has $3K$ fewer parameters than the spatial model,
as there are three $\alpha$ parameters that describe the continuous differentiation effect of distance per cluster,
and they are fixed in the nonspatial model.

@@

@@

\clearpage
\begin{figure}
	\centering
		{\includegraphics[width=4in,height=2.5in]{figs/sims/sim_setup.png}}
		\caption{Schematic of how we simulate datasets with continuous and discrete differentiation, using $K=2$ as an example.  
			    Going forward in time, the $K$ clusters split from a common ancestor at time $\tau_{\text{s}}$,
			    then each expand to colonize a lattice of demes with nearest-neighbor symmetric migration at time $\tau_{\text{e}}$,
			    then finally at time $\tau_{\text{a}}$ collapse into a single lattice consisting of demes 
			    with ancestry entirely in one or the other of the clusters,
			    or admixed between them.
			    }\label{sim_setup}
\end{figure}
